---
title: "Rhizosphere metabolome project - analysis of nutrient data"
author: "Benjamin Delory"
date: "31/08/2020"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    fig_caption: TRUE
---

This document details the R pipeline used to visualize differences in nutrient concentration between soil solution samples collected during the first step (step 1) of the experiment presented in **Delory et al (2020) The rhizosphere metabolome triggers species-specific and context-dependent root responses in later arriving plants**. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Load R packages

```{r Load R packages, warning=FALSE, message=FALSE}
library(ggplot2)
```

# Load data in R

```{r Load data in R, message=FALSE, warning=FALSE}
data <- read.csv("C:/Users/Delory/Documents/GitHub/rhizometabolome/data/Data_nutrients.csv", header = TRUE, stringsAsFactors = FALSE)

data$When[data$When=="Before"]<-"BEFORE nutrient addition"
data$When[data$When=="After"]<-"AFTER nutrient addition"

data$Treatment[data$Treatment=="Control"]<-"CTL"

data$When<-factor(data$When, levels=c("BEFORE nutrient addition", "AFTER nutrient addition"))
```

# Plot results

```{r Plot results, fig.height=3.54, fig.width=5.51, message=FALSE, warning=FALSE}
cbPalette <- c("#11264B", "#C9B75E", "#0C6C67")
cex.axis<-9
cex.title<-10

p<-ggplot(data, aes(x=Element, y=Concentration, fill=Treatment))+
        geom_bar(position="dodge2", stat="identity")+
        theme_bw()+
        facet_wrap(~When)+
        scale_x_discrete(limits=c("N","P","K","Ca","Mg"),
                         labels=c(expression(NO[3]^"-"),
                                  expression(PO[4]^"3-"),
                                  expression(K^"+"),
                                  expression(Ca^"2+"),
                                  expression(Mg^"2+")))+
        scale_fill_manual(values=cbPalette)+
        scale_y_continuous(limits=c(0,70), breaks=seq(0,70,by=10))+
        xlab("")+
        ylab(expression(paste("Nutrient concentration (mg ", L^-1, ")")))+
        theme(axis.title=element_text(size=cex.title, color="black", face="plain"),
              axis.title.y=element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
              legend.title=element_blank(),
              legend.position="top",
              legend.key.size = unit(0.6,"line"),
              legend.background = element_rect(fill=NA),
              legend.text=element_text(color="black"),
              axis.text=element_text(size=cex.axis, color="black"),
              strip.text = element_text(size=10, color="black"),
              strip.background=element_rect(color="black"),
              panel.grid=element_blank())

p

#ggsave(p, filename = "Nutrients_exudates.tiff", dpi=600, compression="lzw", width=14, height=9, units="cm")
```

